### 2019-01-04 documentation/paper.tex

\documentclass[11pt,a4paper]{scrartcl}
%{{{ general stuff
\usepackage[a4paper,bindingoffset=0.2in,%
            left=1in,right=1in,top=1in,bottom=1in,%
            footskip=.25in]{geometry}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{cite}
\usepackage[hidelinks]{hyperref}
\usepackage{float} % use H in figure placement
\pagenumbering{gobble}

%}}}
%{{{ graphics
\usepackage{graphicx} % Bilder
\usepackage{tikz, pgfplots}
\pgfplotsset{compat=1.14}
%}}}
%{{{ math
\usepackage{mathrsfs} % mathcal and mathscr
\usepackage{mathtools, amssymb, amsthm}
\usepackage{bm} % cool bold symbols
%}}}
% a todo command===============================================================
\usepackage[ruled]{algorithm2e}
\newcounter{todocounter}
\newcommand{\todo}[2][noisnotdefined]{
 \marginpar{\fcolorbox{black}{orange}{\footnotesize\textbf{todo}}
 \ifthenelse{\equal{#1}{noisnotdefined}}{}{\textcolor{black}{\newline\tiny #1}}}
 \textbf{\ifthenelse{\equal{#2}{.}}
   {\fcolorbox{orange}{white}{\textcolor{orange}{$\maltese$}}}{{\textcolor{orange}{#2}}}}
 \refstepcounter{todocounter}}

%===============================================================================

\date{}
\title{todo: think about good title}
\author{Felix Bartel and D\'avid Kerekes}

\begin{document}

\maketitle

\section{Introduction}

\todo{include reference \cite{Garis90geneticprogramming} (beginning of neuroevolution, cited by Sims)}
\todo{include reference \cite{Sims94evolving3d} (Sims himself)}

We want to apply a genetic algorithm to a neural network which is known as neuroevolution.
A big advantage of this scheme is that one only needs an indicator how good the neural network performs given a certain task.
This can for instance be given by the outcome of a game.
Our game of choice was Blobby Volley 2 which is a simple two-dimensional volleyball game with open source written in C++ with lua support, see Figure \ref{fig:screenshot}.

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/screenshot.png}
\caption{Screenshot of the game.}
\label{fig:screenshot}
\end{figure}

A scene from the game could be described by the $x$- and $y$-coordinates of the players and the ball, their corresponding velocities and the current number of ball contacts.
These arguments or a subset of them could then be used as input for the neural network.
The output would then be the players movement which only consists of going right, left or jumping.
Assuming the layers of the neural network have $n_1,\dots,n_L$ nodes we then have to optimize \[N = \sum_{l=1}^{L-1} (n_l+1) \, n_{l+1}\] number of variables if we fix the size of the network.

The genotype-representation of the network is accomplished through python objects. This will allow a straightforward implementation of a wide variety of genetic operators, while having next to no impact on the speed of training.
As the evaluation of the population is time-consuming, we do this in a thread-parallel fashion, allowing us to exploit multiple high core-count machines at our disposal.

\begin{figure}[H]
\begin{tabular}{c|ccccccc}
vs. & 
\texttt{ Axji-0-2 } &
\texttt{ COM\_11 } &
\texttt{ GintonicV9 } &
\texttt{ Hypo14 } &
\texttt{ reduced } &
\texttt{ Union } &
\texttt{ trainer } \\
\hline
\texttt{ Axji-0-2 } & - & 0 & 0 & 4 & 0 & 0 & 10 \\
\texttt{ COM\_11 } & 21 & - & 9 & 19 & 6 & 16 & 20 \\
\texttt{ GintonicV9 } & 21 & 12 & - & 19 & 6 & 16 & 20 \\
\texttt{ Hypo14 } & 17 & 2 & 2 & - & 2 & 5 & 15 \\
\texttt{ reduced } & 21 & 19 & 15 & 19 & - & 19 & 21 \\
\texttt{ Union } & 21 & 14 & 5 & 16 & 2 & - & 20 \\
\texttt{ trainer } & 11 & 0 & 1 & 6 & 0 & 1 & -
\end{tabular}
\caption{scores of bots competing against each other}
\end{figure}

\todo{explain the table}

\section{Standard genetic algorithm}

fixed the size of the neural network to $6$ input nodes, one hidden layer with $7$ nodes and $2$ output nodes.

\subsection{Crossover}

As crossover we randomly mixed the nodes with belonging weights and biases of two networks and chose the parents via the fitness proportional roulette wheel selection.

\subsection{Mutation}

Mutation is accomplished by adding Gaussian noise to a subset of all weights and biases.

\subsection{scheme}

Furthermore we implemented elitism.

fitness is the normalized point difference after $21$ serves.

We implemented our own bot to act as an initial trainer \texttt{trainer} for the network.
This trainer is easier to defeat.
We made a copy of each bot where it gives up after a certain amount of touches with the ball, such that normal gameplay will be rewarded.
The simplification gives the evolutionary algorithm the opportunity to start optimizing the neural network in a valid direction, without introducing additional factors to the fitness score.

\subsection{Numerical tests}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced} & \texttt{com\_11} & \texttt{gintonicV9} \\
\includegraphics[width=0.3\textwidth]{img/standard_reduced.png} &
\includegraphics[width=0.3\textwidth]{img/standard_com_11.png} &
\includegraphics[width=0.3\textwidth]{img/standard_gintonicV9.png} \\
\texttt{hyp014} & \texttt{Union} & \texttt{trainer} \\
\includegraphics[width=0.3\textwidth]{img/standard_hyp014.png} &
\includegraphics[width=0.3\textwidth]{img/standard_Union.png} &
\includegraphics[width=0.3\textwidth]{img/standard_trainer.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:standard}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/standard.png}
\caption{trained bots against all other}
\label{fig:standard}
\end{figure}

\todo{coclusion: bots too hard we therefore reward gameplay by the *\_5t bots}

\todo{introduce the goal of an overall good bot and explain the plot along the way (added fitnesses (score of 12 would mean winneing every point against every bot))}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced\_5t} & \texttt{com\_11\_5t} & \texttt{gintonicV9\_5t} \\
\includegraphics[width=0.3\textwidth]{img/standard_reduced_5t.png} &
\includegraphics[width=0.3\textwidth]{img/standard_com_11_5t.png} &
\includegraphics[width=0.3\textwidth]{img/standard_gintonicV9_5t.png} \\
\texttt{hyp014\_5t} & \texttt{Union\_5t} & \texttt{trainer\_5t} \\
\includegraphics[width=0.3\textwidth]{img/standard_hyp014_5t.png} &
\includegraphics[width=0.3\textwidth]{img/standard_Union_5t.png} &
\includegraphics[width=0.3\textwidth]{img/standard_trainer_5t.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:standard_5t}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/standard_5t.png}
\caption{trained bots against all other}
\label{fig:standard_5t}
\end{figure}

\todo{conclusion: better, and not only for 5t bots (algorithm has something to grab on)}

\todo{overall: not very pretty graphs -> need self adaption}

\section{Self adapt 1}

\todo{explain what we are doing}

\subsection{Numerical tests}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced} & \texttt{com\_11} & \texttt{gintonicV9} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_reduced.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_com_11.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_gintonicV9.png} \\
\texttt{hyp014} & \texttt{Union} & \texttt{trainer} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_hyp014.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_Union.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_trainer.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:self_adapt_1}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/self_adapt_1.png}
\caption{trained bots against all other}
\label{fig:self_adapt_1}
\end{figure}

\todo{we can see that the algorithm is converging}
\todo{because trainer is beatable, the nn has something to grab on and performs way better thatn the others}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced\_5t} & \texttt{com\_11\_5t} & \texttt{gintonicV9\_5t} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_reduced_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_com_11_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_gintonicV9_5t.png} \\
\texttt{hyp014\_5t} & \texttt{Union\_5t} & \texttt{trainer\_5t} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_hyp014_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_Union_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_trainer_5t.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:self_adapt_1_5t}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/self_adapt_2_5t.png}
\caption{trained bots against all other}
\label{fig:self_adapt_2_5t}
\end{figure}

\todo{now it's pretty}
\todo{trainer\_5t bad because it is very easy to beat and therefore we can get loast in a local optimum real quick}

\section{Self adapt 2}

\subsection{Numerical tests}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced} & \texttt{com\_11} & \texttt{gintonicV9} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_reduced.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_com_11.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_gintonicV9.png} \\
\texttt{hyp014} & \texttt{Union} & \texttt{trainer} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_hyp014.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_Union.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_trainer.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:self_adapt_2}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/self_adapt_2.png}
\caption{trained bots against all other}
\label{fig:self_adapt_2}
\end{figure}

\todo{similar to section before without 5t}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced\_5t} & \texttt{com\_11\_5t} & \texttt{gintonicV9\_5t} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_reduced_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_com_11_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_gintonicV9_5t.png} \\
\texttt{hyp014\_5t} & \texttt{Union\_5t} & \texttt{trainer\_5t} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_hyp014_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_Union_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_trainer_5t.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:self_adapt_2_5t}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/self_adapt_2_5t.png}
\caption{trained bots against all other}
\label{fig:self_adapt_2_5t}
\end{figure}

\todo{not very predictable behaviour}
\todo{works good for a few examples}
\todo{remember prof has never seen this stuff working and will be happy if it has disadvantages}

\section{Changing opponent}

200 gen \texttt{gintonicV9\_5t}, 200 gen \texttt{com\_11\_5t}, 200 gen \texttt{gintonicV9}, 200 gen \texttt{com\_11}.
\todo{argument with the previous plots that this is a good idea}
\todo{other ideas?}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
standard & self adapt 1 & self adapt 2 \\
\includegraphics[width=0.3\textwidth]{img/standard_gintonicV9_5t_com_11_5t_gintonicV9_com_11.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_gintonicV9_5t_com_11_5t_gintonicV9_com_11.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_gintonicV9_5t_com_11_5t_gintonicV9_com_11.png} \\
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:changing opponent}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/change.png}
\caption{trained bots against all other}
\label{fig:change}
\end{figure}

\todo{only self adapt 1 works}
\todo{we could do more with more generations but loose the comparebility to the previous section}

\section{Self-training}

\subsection{last best}

\begin{figure}[H]
\center
\includegraphics[width=0.155555\textwidth]{img/scheme_self_training_1.png}
\caption{Scheme for self training variant 1}
\label{fig:scheme_self_training_1}
\end{figure}

Fitness evaluated with the best bot of the last generation.

\begin{figure}[H]
\center
\begin{tabular}{cc}
standard & self adapt 1\\
\includegraphics[width=0.3\textwidth]{img/standard_last_best.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_last_best.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:last_best}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.3\textwidth]{img/self_training_1.png}
\caption{trained bots against all other}
\label{fig:self_training_1}
\end{figure}

\todo{risk to end up in local optimum like self adapt 1?}

\subsection{two species}

\begin{figure}[H]
\center
\includegraphics[width=0.155555\textwidth]{img/scheme_self_training_1.png}
\caption{Scheme for self training variant 1}
\label{fig:scheme_self_training_2}
\end{figure}

\begin{figure}[H]
\center
\begin{tabular}{cc}
left side & right side \\
\includegraphics[width=0.3\textwidth]{img/standard_self_training_2_1.png} &
\includegraphics[width=0.3\textwidth]{img/standard_self_training_2_2.png} \\
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum standard mutation.}
\label{fig:standard_self_training_2}
\end{figure}

\todo{too much diversity}

\begin{figure}[H]
\center
\begin{tabular}{cc}
left side & right side \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_self_training_2_1_oldrun.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_self_training_2_2_oldrun.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum with self adapt 1.}
\label{fig:self_adapt_1_self_training_2_oldrun}
\end{figure}

\todo{sadly they do not change too often}
\todo{there is still improvement even though it can not be seen in the plots because they improve themselves}

\begin{figure}[H]
\center
\begin{tabular}{cc}
left side & right side \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_self_training_2_1.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_self_training_2_2.png} \\
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum with self adapt 1 and more generations.}
\label{fig:self_adapt1_self_training_2}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/self_training_2.png}
\caption{trained bots against all other}
\label{fig:self_training_2}
\end{figure}

\todo{we got a pretty awesome bot there which gets beaten only by reduced but never trained against it}

\section{Discussion and Future Ideas}

\todo{everything we have not done}
\todo{train the self trained bot with reduced so we maybe get a perfect score of 12}
\todo{we have not used too many inputs I still coudn't beat the bot which trained itself for four hours}
\todo{acknowledgement to the compute server}

\bibliography{refs}{}
\bibliographystyle{plain}

\end{document}