### 2019-01-07 documentation/paper.tex
Holy hell it's late

\documentclass[11pt,a4paper]{scrartcl}
%{{{ general stuff
\usepackage[a4paper,bindingoffset=0.2in,%
            left=1in,right=1in,top=1in,bottom=1in,%
            footskip=.25in]{geometry}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{cite}
\usepackage[hidelinks]{hyperref}
\usepackage{float} % use H in figure placement
% \pagenumbering{gobble}

%}}}
%{{{ graphics
\usepackage{graphicx} % Bilder
\usepackage{tikz, pgfplots}
\pgfplotsset{compat=1.14}
%}}}
%{{{ math
\usepackage{mathrsfs} % mathcal and mathscr
\usepackage{mathtools, amssymb, amsthm}
\usepackage{bm} % cool bold symbols
%}}}
% a todo command===============================================================
\usepackage[ruled]{algorithm2e}
\newcounter{todocounter}
\newcommand{\todo}[2][noisnotdefined]{
 \marginpar{\fcolorbox{black}{orange}{\footnotesize\textbf{todo}}
 \ifthenelse{\equal{#1}{noisnotdefined}}{}{\textcolor{black}{\newline\tiny #1}}}
 \textbf{\ifthenelse{\equal{#2}{.}}
   {\fcolorbox{orange}{white}{\textcolor{orange}{$\maltese$}}}{{\textcolor{orange}{#2}}}}
 \refstepcounter{todocounter}}

%===============================================================================

\date{\today}
\title{Really Good Title}
\subtitle{Genetic Algorithms and Evolutionary Computing Project}
\author{Felix Bartel and D\'avid Kerekes}

\begin{document}

\maketitle
\todo{think about good title}

\section{Introduction}

Undoubtedly, the neural network architecture has been the star of the machine learning resurgence of the past few years. It's provable quality of being an universal function approximator makes it a sensible choice for various applications. The main algorithm used for training, backpropagation however is subject to an array of criticism from various authors \cite{some_papers_about_backprop}, and there are various attempts at improving or replacing it \cite{that_paper_where_ga_beats_backprop}.
\todo{Add a sentence of context for \cite{Sims94evolving3d} and \cite{Garis90geneticprogramming}}

In this project we explore some properties of neuroevolution, the technique of applying a genetic algorithm to a neural network \cite{Garis90geneticprogramming} \cite{Sims94evolving3d}.
A big advantage of this scheme is that for a series of different inputs a single scalar indicator of performance is sufficient in order to optimize a neural network for a certain task.
This can for instance be given by the outcome of a game.
Our game of choice is Blobby Volley 2 which is an open source, simple two-dimensional volleyball game written in C++ with lua support. See Figure \ref{fig:screenshot}.

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/screenshot.png}
\caption{Screenshot of the game.}
\label{fig:screenshot}
\end{figure}

\section{Training environment}

A scene from the game could be described by the $x$- and $y$-coordinates of the players and the ball, their corresponding velocities and the current number of ball contacts.
These arguments or a subset of them can then be used as input for the neural network.
The output should correspond to the possible movement options of the player, which consists of going right, going left, or jumping.
Assuming the layers of the neural network have $n_1,\dots,n_L$ nodes we then have to optimize \(N = \sum_{l=1}^{L-1} (n_l+1) \, n_{l+1}\) number of variables.

The game comes bundled with a number of bots with different styles, that are designed to pose a challenge to a wide range of human opponents. In order to define a clean ranking between them, we recorded the score of all possible lineups. The results can be seen in table no. \ref{tab:bot_vs_bot}.
\todo{explain the table}

\begin{figure}[H]
\begin{tabular}{c|ccccccc}
vs. &
\texttt{Axji-0-2} &
\texttt{COM\_11} &
\texttt{GintonicV9} &
\texttt{Hypo14} &
\texttt{reduced} &
\texttt{Union} &
\texttt{trainer} \\
\hline
\texttt{Axji-0-2} & - & 0 & 0 & 4 & 0 & 0 & 10 \\
\texttt{COM\_11} & 21 & - & 9 & 19 & 6 & 16 & 20 \\
\texttt{GintonicV9} & 21 & 12 & - & 19 & 6 & 16 & 20 \\
\texttt{Hypo14} & 17 & 2 & 2 & - & 2 & 5 & 15 \\
\texttt{reduced} & 21 & 19 & 15 & 19 & - & 19 & 21 \\
\texttt{Union} & 21 & 14 & 5 & 16 & 2 & - & 20 \\
\texttt{trainer} & 11 & 0 & 1 & 6 & 0 & 1 & -
\end{tabular}
\label{tab:bot_vs_bot}
\caption{Scores of bots competing against each other}
\end{figure}

We implemented our own bot to act as an initial trainer \texttt{trainer} for the network.
This trainer is easier to defeat.
We made a copy of each bot where it gives up after a certain amount of touches with the ball, such that normal gameplay will be rewarded.
The simplification gives the evolutionary algorithm the opportunity to start optimizing the neural network in a valid direction, without introducing additional factors to the fitness score.

\section{Implementation}
The genotype-representation of the network is accomplished through python objects. This allows a straightforward implementation of a wide variety of genetic operators. Since our algorithm is constrained overwhelmingly by evaluation time, these inefficient but easy to modify structures do not impact performance.
As the evaluation of the population is time-consuming, we do this in a thread-parallel fashion, allowing us to exploit multiple high core-count machines at our disposal. In order to achieve this, we modified the original game by replacing the graphics and sound modules with dummies.

\section{Initial genetic algorithm}
Here we describe our starting point for the genetic algorithm. All subsequent versions are based off of this variant, with not mentioned details staying the same.

\subsection{Network architecture}

For the whole experiment, we fixed the size of the neural network to $6$ input nodes, one hidden layer with $7$ nodes and $2$ output nodes. As can be seen later, this proved sufficient for a surprisingly high performance by the network. The input nodes got fed with the x and y positions of the player, it's relative position compared to the ball, and the velocity vector of the ball.
The output nodes consisted of two neurons with tanh activations, one responsible for going left or right in the next timestep (if the output was below or above of threshold values $t_l = 0.49, t_r = 0.51$) and the other for jumping (above threshold $t_j = 0.7$).

\subsection{Scheme}
The algorithm worked with a population size of 100, with 150 offspring after crossover. Our survivor selection strategy was elitism, selecting from the pool of children plus the 10 best parents from the last generation.

The fitness function was the normalized point difference after $21$ serves.

\subsection{Mutation}
Mutation is accomplished by adding Gaussian noise to a subset of all weights and biases. We selected an individual for mutation with a probability of $0.7$, which then had a per-neuron mutation rate of $0.5$, and a sigma of $0.01$.

\subsection{Crossover}
As crossover we randomly mixed the nodes with belonging weights and biases of two networks and chose the parents via the fitness proportional roulette wheel selection.

\subsection{Numerical tests}
We trained six networks, each against a different bot. We test the generalization of both the training method and the best specimens by evaluating all the networks against all the bots. A score for each network is calculated by summing up it's fitness value for all cases (see Fig. \ref{fig:standard_c}), so a network consistently beating all opponents would receive a score of 12. We expect that a good training method would produce networks that all beat their respective trainers and also fare well against other opponents. The chances of happening upon a network that can exploit a trainer by "accident" (but is susceptible to abuse from a human player for example) decrease exponentially by evaluating on multiple opponents.

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced} & \texttt{com\_11} & \texttt{gintonicV9} \\
\includegraphics[width=0.3\textwidth]{img/standard_reduced.png} &
\includegraphics[width=0.3\textwidth]{img/standard_com_11.png} &
\includegraphics[width=0.3\textwidth]{img/standard_gintonicV9.png} \\
\texttt{hyp014} & \texttt{Union} & \texttt{trainer} \\
\includegraphics[width=0.3\textwidth]{img/standard_hyp014.png} &
\includegraphics[width=0.3\textwidth]{img/standard_Union.png} &
\includegraphics[width=0.3\textwidth]{img/standard_trainer.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum while training against different bots.}
\label{fig:standard}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/standard.png}
\caption{Cumulative sum of the trained neural networks against all the bots.}
\label{fig:standard_c}
\end{figure}

We can see that the algorithm achieves reasonable scores against the two easiest opponents, \texttt{trainer} and \texttt{hyp014}. Since we were not able to get significantly better results by hyperparameter-tweaking, we theorized that the discrepancy might come from the trainers being too hard to beat by any initial network mutation, therefore providing the same zero fitness for any single useful mutation. Therefore we introduced a weaker, "benevolent" set of trainers that give up after scoring five points by not returning the ball anymore. This lets a specimen that can pass initial serves back receive a positive fitness score. These bots are marked by appending the \texttt{\_5t} suffix to their name.

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced\_5t} & \texttt{com\_11\_5t} & \texttt{gintonicV9\_5t} \\
\includegraphics[width=0.3\textwidth]{img/standard_reduced_5t.png} &
\includegraphics[width=0.3\textwidth]{img/standard_com_11_5t.png} &
\includegraphics[width=0.3\textwidth]{img/standard_gintonicV9_5t.png} \\
\texttt{hyp014\_5t} & \texttt{Union\_5t} & \texttt{trainer\_5t} \\
\includegraphics[width=0.3\textwidth]{img/standard_hyp014_5t.png} &
\includegraphics[width=0.3\textwidth]{img/standard_Union_5t.png} &
\includegraphics[width=0.3\textwidth]{img/standard_trainer_5t.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:standard_5t}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/standard_5t.png}
\caption{Trained networks against all other bots.}
\label{fig:standard_5t_aa}
\end{figure}

As Fig. \ref{fig:standard_5t} and \ref{fig:standard_5t_aa} show, the networks evolved against the benevolent trainers give significantly better performance against the whole batch of bots, reaching twice the maximum cumulative score.

\section{Per-individual self adaptation}

\todo{overall: not very pretty graphs <<-- Why do we decided to do self adaptation?}

We first implemented a method similar to (4) from \cite{self_adapt}, giving each individual a mutation rate and a sigma value. These were mutated when the individual was selected for mutation. Their initialization was also random, in the range of [0.25, 0.75] for the mutation rate and [0, 0.03] for the sigma values.

\subsection{Numerical tests}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced} & \texttt{com\_11} & \texttt{gintonicV9} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_reduced.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_com_11.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_gintonicV9.png} \\
\texttt{hyp014} & \texttt{Union} & \texttt{trainer} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_hyp014.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_Union.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_trainer.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:self_adapt_1}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/self_adapt_1.png}
\caption{Trained networks against all other bots.}
\label{fig:self_adapt_1_aa}
\end{figure}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced\_5t} & \texttt{com\_11\_5t} & \texttt{gintonicV9\_5t} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_reduced_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_com_11_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_gintonicV9_5t.png} \\
\texttt{hyp014\_5t} & \texttt{Union\_5t} & \texttt{trainer\_5t} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_hyp014_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_Union_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_trainer_5t.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:self_adapt_1_5t}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/self_adapt_2_5t.png}
\caption{Trained networks against all other bots.}
\label{fig:self_adapt_1_5t_aa}
\end{figure}

We can see in figures \ref{self_adapt_1_5t} and \ref{self_adapt_1_5t_aa} that the results improve significantly in both cases. Self adaptation causes the networks trained on \texttt{_5t} opponents to get consistent results across the board.

\todo{trainer\_5t does not coincide on the two graphs, what happened? We get a good result against trainer_5t according to the bottom graph}

\section{Per-parameter self adaptation}

We also explored a method similar to (5) in \cite{self_adapt}, where each parameter gets it's own mutation sigma parameter. This produces similar results to the harsh teachers with per-individual self adaptation

\subsection{Numerical tests}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced} & \texttt{com\_11} & \texttt{gintonicV9} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_reduced.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_com_11.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_gintonicV9.png} \\
\texttt{hyp014} & \texttt{Union} & \texttt{trainer} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_hyp014.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_Union.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_trainer.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:self_adapt_2}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/self_adapt_2.png}
\caption{Trained networks against all other bots.}
\label{fig:self_adapt_2_aa}
\end{figure}

\todo{similar to section before without 5t}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
\texttt{reduced\_5t} & \texttt{com\_11\_5t} & \texttt{gintonicV9\_5t} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_reduced_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_com_11_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_gintonicV9_5t.png} \\
\texttt{hyp014\_5t} & \texttt{Union\_5t} & \texttt{trainer\_5t} \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_hyp014_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_Union_5t.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_trainer_5t.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:self_adapt_2_5t}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/self_adapt_2_5t.png}
\caption{Trained networks against all other bots.}
\label{fig:self_adapt_2_5t_aa}
\end{figure}

\todo{not very predictable behaviour}
\todo{works good for a few examples}
\todo{remember prof has never seen this stuff working and will be happy if it has disadvantages}

\section{Changing opponent}

200 gen \texttt{gintonicV9\_5t}, 200 gen \texttt{com\_11\_5t}, 200 gen \texttt{gintonicV9}, 200 gen \texttt{com\_11}.
\todo{argument with the previous plots that this is a good idea}
\todo{other ideas?}

\begin{figure}[H]
\center
\begin{tabular}{ccc}
standard & self adapt 1 & self adapt 2 \\
\includegraphics[width=0.3\textwidth]{img/standard_gintonicV9_5t_com_11_5t_gintonicV9_com_11.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_gintonicV9_5t_com_11_5t_gintonicV9_com_11.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_2_gintonicV9_5t_com_11_5t_gintonicV9_com_11.png} \\
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:changing opponent}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/change.png}
\caption{Trained networks against all other bots.}
\label{fig:change}
\end{figure}

\todo{only self adapt 1 works}
\todo{we could do more with more generations but loose the comparebility to the previous section}

\section{Self-training}

\subsection{last best}

\begin{figure}[H]
\center
\includegraphics[width=0.155555\textwidth]{img/scheme_self_training_1.png}
\caption{Scheme for self training variant 1}
\label{fig:scheme_self_training_1}
\end{figure}

Fitness evaluated with the best bot of the last generation.

\begin{figure}[H]
\center
\begin{tabular}{cc}
standard & self adapt 1\\
\includegraphics[width=0.3\textwidth]{img/standard_last_best.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_last_best.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum.}
\label{fig:last_best}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.3\textwidth]{img/self_training_1.png}
\caption{Trained networks against all other bots.}
\label{fig:self_training_1}
\end{figure}

\todo{risk to end up in local optimum like self adapt 1?}

\subsection{two species}

\begin{figure}[H]
\center
\includegraphics[width=0.155555\textwidth]{img/scheme_self_training_1.png}
\caption{Scheme for self training variant 1}
\label{fig:scheme_self_training_2}
\end{figure}

\begin{figure}[H]
\center
\begin{tabular}{cc}
left side & right side \\
\includegraphics[width=0.3\textwidth]{img/standard_self_training_2_1.png} &
\includegraphics[width=0.3\textwidth]{img/standard_self_training_2_2.png} \\
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum standard mutation.}
\label{fig:standard_self_training_2}
\end{figure}

\todo{too much diversity}

\begin{figure}[H]
\center
\begin{tabular}{cc}
left side & right side \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_self_training_2_1_oldrun.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_self_training_2_2_oldrun.png}
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum with self adapt 1.}
\label{fig:self_adapt_1_self_training_2_oldrun}
\end{figure}

\todo{sadly they do not change too often}
\todo{there is still improvement even though it can not be seen in the plots because they improve themselves}

\begin{figure}[H]
\center
\begin{tabular}{cc}
left side & right side \\
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_self_training_2_1.png} &
\includegraphics[width=0.3\textwidth]{img/self_adapt_1_self_training_2_2.png} \\
\end{tabular}
\caption{Fitness density with indicated minimum, average and maximum with self adapt 1 and more generations.}
\label{fig:self_adapt1_self_training_2}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=0.4\textwidth]{img/self_training_2.png}
\caption{Trained networks against all other bots.}
\label{fig:self_training_2}
\end{figure}

\todo{we got a pretty awesome bot there which gets beaten only by reduced but never trained against it}

\section{Discussion and Future Ideas}

\todo{rationalize not messing around with more mutation and crossover types as we promised}

\todo{everything we have not done}
\todo{train the self trained bot with reduced so we maybe get a perfect score of 12}
\todo{we have not used too many inputs I still coudn't beat the bot which trained itself for four hours}
\todo{acknowledgement to the compute server}

\bibliography{refs}{}
\bibliographystyle{plain}

\end{document}